Between Subgraph Isomorphism and Maximum Common Subgraph

### 2016-09-11 paper/paper.tex
Committing from 25,000ft

@@

\subsection{Experimental Setup}

@@

Note that many of these instances are \emph{much} larger than were used in the recent comparison of
maximum common subgraph algorithms by \citet{DBLP:conf/cp/McCreeshNPS16}: for unlabelled graphs,
they used pairs of graphs with up to 35 vertices in each, whilst our dataset contains patterns with
up to ?? vertices and targets with up to ?? vertices. This causes problems: both the maximum common
subgraph algorithm and the clique encoding require $O(\left|\V(P)\right|^2\left|\V(T)\right|^2)$
memory. For maximum common subgraph, 1,560 of the instances cannot fit in the amount of RAM we have
available---we treat these instances as having timed out. The situation for the clique encoding is
even worse, and 3,653 of these instances do not fit in RAM. In contrast, the approach we discuss
here requires only $O(\left|\V(P)\right|^2\left|\V(T)\right|)$ space (which would still be a problem
on some desktop machines, but not on our server hardware).

@@

A non-induced $k$-less-subgraph isomorphism from $P$ to $T$ is equivalent to a subgraph
isomorphism between $P$ and $T$ with $k$ extra universally-adjacent vertices. (However, we can do a
bit better algorithmically.) For induced $k$-less-subgraph isomorphisms, such an approach cannot
work.

@@

Unlike in conventional subgraph isomorphism, we cannot extend this filtering to look at paths of
length three: removing a single vertex can delete arbitrarily many such paths. We could count paths
of any length which are vertex disjoint, although calculating this is too expensive to be useful in
practice.

To allow for fast propagation, rather than calculating paths dynamically, we follow the approach of
\citet{DBLP:conf/cp/McCreeshP15} and construct supplemental graphs. We use paths of length 2,
looking at counts of 1, 2, and 3 in the target graph (and so counts of $1 + k$ up to $3 + k$ in the
pattern).  We then investigate whether this leads to new constraints being generated.

@@

\section{A New Algorithm}

@@

\section{Empirical Evaluation}

>>>>

\subsection{Algorithms for the $k$-Less Problem}

A non-induced $k$-less-subgraph isomorphism from $P$ to $T$ is equivalent to a subgraph isomorphism
between $P$ and $T$ with $k$ extra universally-adjacent vertices added to $T$. However, using such
an approach is not ideal, because it would introduce symmetries; for induced $k$-less-subgraph
isomorphisms, an approach based around adding vertices to $T$ cannot work at all.

Another algorithmic approach could be to try each way of removing $k$ vertices from the pattern
graph, and solving the subgraph isomorphism problem in turn. This approach might be feasible for $k
= 1$, although it would involve a lot of duplication of search effort, but for larger values of $k$
the number of searches which would have to be made would grow very quickly.

Finally, for the induced case we could try adapting maximum common subgraph algorithms to solve the
decision problem. However, both the maximum common subgraph algorithm and the clique encoding
require $O(\left|\V(P)\right|^2\left|\V(T)\right|^2)$ memory, which is extremely problematic on the
instances we wish to work with.

Instead, we introduce a new algorithm, inspired by a state of the art subgraph isomorphism
algorithm. The approach we discuss requires only $O(\left|\V(P)\right|^2\left|\V(T)\right|)$ space
(this can be thousands of times less than maximum common subgraph approaches, once the constant
factors and orders of $T$ in our instances are considered). We list the algorithm in
\cref{algorithm:sip}, explain it fully in \cref{section:algorithm}, and evaluate it in
\cref{section:evaluation}, but first we will discuss our experimental setup, and then detail some
experiments to verify that we can gain effective domain filtering (\cref{section:degreefiltering})
and implied constraints (\cref{section:pathfiltering}) by using subgraph isomorphism style reasoning
about degrees and paths in the $k$-less setting.

@@

\section{Experimental Setup}

@@

Note that many of these instances are \emph{much} larger than were used in the recent comparison of
maximum common subgraph algorithms by \citet{DBLP:conf/cp/McCreeshNPS16}: for unlabelled graphs,
they used pairs of graphs with up to 35 vertices in each, whilst our dataset contains patterns with
up to ?? vertices and targets with up to ?? vertices. For maximum common subgraph, 1,560 of the
instances cannot fit in the amount of RAM we have available---we treat these instances as having
timed out. The situation for the clique encoding is even worse, and 3,653 of these instances do not
fit in RAM. In contrast, for our algorithm, every instance fits comfortably (at least on modern
server-class hardware).

@@

@@

Unlike in conventional subgraph isomorphism, we cannot extend this filtering to look at paths of
length three: removing a single vertex can delete arbitrarily many such paths. We \emph{could}
instead count paths of any length which are vertex disjoint, although calculating this appears to be
too expensive to be beneficial in practice.

To allow for fast propagation, rather than calculating paths dynamically, we follow the approach of
\citet{DBLP:conf/cp/McCreeshP15} and construct supplemental graphs, and find a mapping which is
simultaneously a non-induced subgraph isomorphism between every supplemental graph pair. We use
paths of length 2, looking at counts of 1, 2, and 3 in the target graph (and so counts of $1 + k$ up
to $3 + k$ in the pattern).  We then investigate whether this leads to new constraints being
generated.

@@

\section{A New Algorithm}\label{section:algorithm}

@@

\section{Empirical Evaluation}\label{section:evaluation}
